---
title: "LiverGwena"
author: "MAH"
date: "6/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Install Essential Libraries
```{r}

if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

if (!requireNamespace("tidyverse", quietly = TRUE))
    install.packages("tidyverse")

if (!requireNamespace("data.table", quietly = TRUE))
    install.packages("data.table")

if (!requireNamespace("corrplot", quietly = TRUE))
    install.packages("corrplot")

if (!requireNamespace("GWENA", quietly = TRUE))
BiocManager::install("GWENA")

if (!requireNamespace("enrichR", quietly = TRUE))
    install.packages("enrichR")

if (!requireNamespace("devtools", quietly = TRUE))
    install.packages("devtools")

if (!requireNamespace("factoextra", quietly = TRUE))
    install.packages("factoextra")

if (!requireNamespace("FactoMineR", quietly = TRUE))
    install.packages("FactoMineR")

if (!requireNamespace("here", quietly = TRUE))
    install.packages("here")
```


Load essential libraries

```{r}

library("GWENA")
library("tidyverse")
library("data.table")
library("here")
library("corrplot")

library("devtools")
library("FactoMineR")
library("factoextra")
library("Biobase")
library("enrichR")

library("readxl")
library("openxlsx")

```


Transpose data so that each column is a gene and each row is a sample/observation


```{r}

LiverGenesOriginal <- read.table(here("Results", "Liver", "Consolidated_UnqRefseq_Cpm_AllF2_d42.txt"), header = TRUE)
LiverGenesOriginal[1:10,1:10]

rownames(LiverGenesOriginal) <- LiverGenesOriginal$gene
LiverGenesOriginal$gene <- NULL
LiverGenesOriginal[1:10,1:10]

LiverGenesTrans <- as.matrix(LiverGenesOriginal)
LiverGenesTrans <- t(LiverGenesTrans)
LiverGenesTrans <- as.data.frame(LiverGenesTrans)
LiverGenesTrans[1:10,1:10]
is_data_expr(LiverGenesTrans)


PhenoData <- read.table(here("Data", "PhenotypeData", "Pheno_data.txt"), na.strings = "NA", header = TRUE)
PhenoData <- PhenoData %>% filter(ID != "s42")

rownames(PhenoData) <- PhenoData$ID
PhenoData$ID <- NULL
PhenoData[1:5,]

```


```{r}

# refers to number of cores to be used, questions whether Rstudio can use more than 1
threads_to_use <- 1 


LiverGenesFiltered <- filter_RNA_seq(LiverGenesTrans, min_count = 10, method = c("at least one"))
LiverGenesFiltered2 <- filter_low_var(LiverGenesFiltered, pct = 0.5, type = "median")

#LiverGenesFiltered <- LiverGenesFiltered[, 1:1000]  # reduce size for testing purposes
#LiverGenesFiltered[1:10,1:10]
#PhenoData[1:5,]

net <- build_net(LiverGenesFiltered2, cor_func = "spearman", n_threads = threads_to_use)

# Power selected :
net$metadata$power

# Fit of the power law to data ($R^2$) :
fit_power_table <- net$metadata$fit_power_table
fit_power_table[fit_power_table$Power == net$metadata$power, "SFT.R.sq"]


```






```{r}

modules <- detect_modules(LiverGenesFiltered2, 
                            net$network, 
                            detailled_result = TRUE,
                            merge_threshold = 0.25)


# Number of modules before merging :
length(unique(modules$modules_premerge))


# Number of modules after merging: 
length(unique(modules$modules))


layout_mod_merge <- plot_modules_merge(
  modules_premerge = modules$modules_premerge, 
  modules_merged = modules$modules)


```





```{r}

ggplot2::ggplot(data.frame(modules$modules %>% stack), 
                ggplot2::aes(x = ind)) + ggplot2::stat_count() +
  ggplot2::ylab("Number of genes") +
  ggplot2::xlab("Module")

```





```{r}

phenotype_association <- associate_phenotype(
  modules$modules_eigengenes, 
  PhenoData %>% dplyr::select(BW, TG_liver, Chol_liver, Glucose_S, Insulin_S, Trig_S, Chol_S, FGF21_S, UricAcid_S))

plot_modules_phenotype(phenotype_association)

```





```{r}

assoc <- t(as.matrix(phenotype_association$association))
pval <- t(as.matrix(phenotype_association$pval))

corrplot(assoc, p.mat = pval, insig = 'label_sig', pch.cex = 1)

pdf(here("Results", "Liver", "GwenaLiverPhenotypeAssociationPlot.pdf"))
corrplot(assoc, p.mat = pval, insig = 'label_sig', pch.cex = 1)
dev.off()


```






```{r}

module0 <- (modules$modules[1])
fwrite(module0,here("Results", "Liver", "module0.txt"), row.names=FALSE, col.names = FALSE)

module1 <- (modules$modules[2])
fwrite(module1,here("Results", "Liver", "module1.txt"), row.names=FALSE, col.names = FALSE)

module2 <- (modules$modules[3])
fwrite(module2,here("Results", "Liver", "module2.txt"), row.names=FALSE, col.names = FALSE)

module3 <- (modules$modules[4])
fwrite(module3,here("Results", "Liver", "module3.txt"), row.names=FALSE, col.names = FALSE)

module4 <- (modules$modules[5])
fwrite(module4, here("Results", "Liver", "module4.txt"), row.names=FALSE, col.names = FALSE)

module5 <- (modules$modules[6])
fwrite(module5,here("Results", "Liver", "module5.txt"), row.names=FALSE, col.names = FALSE)


```





Perform PCA on Modules with significant associations and take first PC to identify weights and on a per gene basis. Separate into gene sets that are positively or negatively associated with the module and use these separately in enrichment analysis in EnrichR.



Analysis for Module 2:

```{r}

module2elem <- module2[[1]]
LiverGenesFiltered3 <- dplyr::select(LiverGenesFiltered2, one_of(module2elem))

ModulePC <- PCA(LiverGenesFiltered3, scale.unit = TRUE, ncp = 1, graph = TRUE)
Eigen <- get_eigenvalue(ModulePC)
fviz_eig(ModulePC)
VariableWeights <- get_pca_var(ModulePC)$coord
VariableWeights <- as.data.frame(VariableWeights)

VariableWeights <- VariableWeights %>% arrange(VariableWeights)

Module2Pos <- VariableWeights %>% filter(VariableWeights >= 0)
Module2Neg <- VariableWeights %>% filter(VariableWeights < 0)

write.table(Module2Pos, file = here("Results", "Liver", "Module2PositiveGenes.txt"), sep="\t", quote=FALSE)
write.table(Module2Neg, file = here("Results", "Liver", "Module2NegativeGenes.txt"), sep="\t", quote=FALSE)

M2Pos <- rownames(Module2Pos)
M2Neg <- rownames(Module2Neg)


setEnrichrSite("Enrichr")
dbs <- c("GO_Biological_Process_2021", "TRRUST_Transcription_Factors_2019")

PosEnriched <- enrichr(M2Pos, dbs)
NegEnriched <- enrichr(M2Neg, dbs)



plotEnrich(PosEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module2PositiveGOBioProcessPlot.pdf"))
plotEnrich(PosEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()


plotEnrich(PosEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module2PositiveTrrustTfPlot.pdf"))
plotEnrich(PosEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()


Posm2Go <- PosEnriched[[1]]
Posm2TF <- PosEnriched[[2]]

plotEnrich(NegEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module2NegativeGOBioProcessPlot.pdf"))
plotEnrich(NegEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()

plotEnrich(NegEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module2NegativeTrrustTfPlot.pdf"))
plotEnrich(NegEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()


Negm2Go <- NegEnriched[[1]]
Negm2TF <- NegEnriched[[2]]

EnrichmentAnalysis <- createWorkbook()

addWorksheet(EnrichmentAnalysis, "PosGOBioProcess2021")
addWorksheet(EnrichmentAnalysis, "NegGOBioProcess2021")
addWorksheet(EnrichmentAnalysis, "PosTRRUST_2019")
addWorksheet(EnrichmentAnalysis, "NegTRRUST_2019")


writeData(EnrichmentAnalysis, sheet = "PosGOBioProcess2021", slice_min(Posm2Go, P.value, n = 20))
writeData(EnrichmentAnalysis, sheet = "NegGOBioProcess2021", slice_min(Negm2Go,P.value, n = 20))

writeData(EnrichmentAnalysis, sheet = "PosTRRUST_2019", slice_min(Posm2TF,P.value, n = 20))
writeData(EnrichmentAnalysis, sheet = "NegTRRUST_2019", slice_min(Negm2TF,P.value, n = 20))

saveWorkbook(EnrichmentAnalysis, here("Results", "Liver", "Module2EnrichmentAnalysis.xlsx"))

```






Analysis for Module 3:

```{r}

module3elem <- module3[[1]]
LiverGenesFiltered3 <- dplyr::select(LiverGenesFiltered2, one_of(module3elem))

ModulePC <- PCA(LiverGenesFiltered3, scale.unit = TRUE, ncp = 1, graph = TRUE)
Eigen <- get_eigenvalue(ModulePC)
fviz_eig(ModulePC)
VariableWeights <- get_pca_var(ModulePC)$coord
VariableWeights <- as.data.frame(VariableWeights)

VariableWeights <- VariableWeights %>% arrange(VariableWeights)

Module3Pos <- VariableWeights %>% filter(VariableWeights >= 0)
Module3Neg <- VariableWeights %>% filter(VariableWeights < 0)

write.table(Module3Pos, file = here("Results", "Liver", "Module3PositiveGenes.txt"), sep="\t", quote=FALSE)
write.table(Module3Neg, file = here("Results", "Liver", "Module3NegativeGenes.txt"), sep="\t", quote=FALSE)

M3Pos <- rownames(Module3Pos)
M3Neg <- rownames(Module3Neg)


setEnrichrSite("Enrichr")
dbs <- c("GO_Biological_Process_2021", "TRRUST_Transcription_Factors_2019")

PosEnriched <- enrichr(M3Pos, dbs)
NegEnriched <- enrichr(M3Neg, dbs)

plotEnrich(PosEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module3PositiveGOBioProcessPlot.pdf"))
plotEnrich(PosEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()


plotEnrich(PosEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module3PositiveTrrustTfPlot.pdf"))
plotEnrich(PosEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()


Posm3Go <- PosEnriched[[1]]
Posm3TF <- PosEnriched[[2]]

plotEnrich(NegEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module3NegativeGOBioProcessPlot.pdf"))
plotEnrich(NegEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()

plotEnrich(NegEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module3NegativeTrrustTfPlot.pdf"))
plotEnrich(NegEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()

Negm3Go <- NegEnriched[[1]]
Negm3TF <- NegEnriched[[2]]


EnrichmentAnalysis <- createWorkbook()

addWorksheet(EnrichmentAnalysis, "PosGOBioProcess2021")
addWorksheet(EnrichmentAnalysis, "NegGOBioProcess2021")
addWorksheet(EnrichmentAnalysis, "PosTRRUST_2019")
addWorksheet(EnrichmentAnalysis, "NegTRRUST_2019")


writeData(EnrichmentAnalysis, sheet = "PosGOBioProcess2021", slice_min(Posm3Go, P.value, n = 20))
writeData(EnrichmentAnalysis, sheet = "NegGOBioProcess2021", slice_min(Negm3Go,P.value, n = 20))

writeData(EnrichmentAnalysis, sheet = "PosTRRUST_2019", slice_min(Posm3TF,P.value, n = 20))
writeData(EnrichmentAnalysis, sheet = "NegTRRUST_2019", slice_min(Negm3TF,P.value, n = 20))

saveWorkbook(EnrichmentAnalysis, here("Results", "Liver", "Module3EnrichmentAnalysis.xlsx"))

```



Analysis for Module 5:

```{r}

module5elem <- module5[[1]]
LiverGenesFiltered3 <- dplyr::select(LiverGenesFiltered2, one_of(module5elem))

ModulePC <- PCA(LiverGenesFiltered3, scale.unit = TRUE, ncp = 1, graph = TRUE)
Eigen <- get_eigenvalue(ModulePC)
fviz_eig(ModulePC)
VariableWeights <- get_pca_var(ModulePC)$coord
VariableWeights <- as.data.frame(VariableWeights)

VariableWeights <- VariableWeights %>% arrange(VariableWeights)

Module5Pos <- VariableWeights %>% filter(VariableWeights >= 0)
Module5Neg <- VariableWeights %>% filter(VariableWeights < 0)

write.table(Module5Pos, file = here("Results", "Liver", "Module5PositiveGenes.txt"), sep="\t", quote=FALSE)
write.table(Module5Neg, file = here("Results", "Liver", "Module5NegativeGenes.txt"), sep="\t", quote=FALSE)

M5Pos <- rownames(Module5Pos)
M5Neg <- rownames(Module5Neg)


setEnrichrSite("Enrichr")
dbs <- c("GO_Biological_Process_2021", "TRRUST_Transcription_Factors_2019")

PosEnriched <- enrichr(M5Pos, dbs)
NegEnriched <- enrichr(M5Neg, dbs)

plotEnrich(PosEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module5PositiveGOBioProcessPlot.pdf"))
plotEnrich(PosEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()


plotEnrich(PosEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module5PositiveTrrustTfPlot.pdf"))
plotEnrich(PosEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()


Posm5Go <- PosEnriched[[1]]
Posm5TF <- PosEnriched[[2]]

plotEnrich(NegEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module5NegativeGOBioProcessPlot.pdf"))
plotEnrich(NegEnriched[[1]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()

plotEnrich(NegEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
pdf(here("Results", "Liver", "Module5NegativeTrrustTfPlot.pdf"))
plotEnrich(NegEnriched[[2]], showTerms = 10, numChar = 75, y = "Ratio", orderBy = "P.value")
dev.off()

Negm5Go <- NegEnriched[[1]]
Negm5TF <- NegEnriched[[2]]


EnrichmentAnalysis <- createWorkbook()

addWorksheet(EnrichmentAnalysis, "PosGOBioProcess2021")
addWorksheet(EnrichmentAnalysis, "NegGOBioProcess2021")
addWorksheet(EnrichmentAnalysis, "PosTRRUST_2019")
addWorksheet(EnrichmentAnalysis, "NegTRRUST_2019")


writeData(EnrichmentAnalysis, sheet = "PosGOBioProcess2021", slice_min(Posm5Go, P.value, n = 20))
writeData(EnrichmentAnalysis, sheet = "NegGOBioProcess2021", slice_min(Negm5Go,P.value, n = 20))

writeData(EnrichmentAnalysis, sheet = "PosTRRUST_2019", slice_min(Posm5TF,P.value, n = 20))
writeData(EnrichmentAnalysis, sheet = "NegTRRUST_2019", slice_min(Negm5TF,P.value, n = 20))

saveWorkbook(EnrichmentAnalysis, here("Results", "Liver", "Module5EnrichmentAnalysis.xlsx"))

```



